\begin{abstract}
    The GDPR's cooperation and consistency mechanism is a network of peer authorities collectively tasked with the uniform application of the law protecting the fundamental right to data protection.
    Since its introduction, the mechanism has been criticized, \emph{inter alia}, for affording the lead supervisory authority too large of a role in setting the scope of investigations, leading to under-investigation and under-enforcement. 
    The recently adopted Regulation 2025/2518 reforms the cooperation and consistency procedure by proceduralizing many aspects of the required cooperation between data protection authorities.
    This study examines the reformed cooperation flow in light of two of the sub-principles of good administration: the duty of care, and the duty to give reasons.
    The reforms are assessed in the context of network theories of peer pressure, looking at whether the proceduralization offered by the Procedural Regulation will effectively curb the dominance of the lead supervisory authority and produce better enforcement outcomes, bolstering the respect of the principle of good administration.
    While the reform brings improvements by providing formal steps for interaction, it does not address the structural issues of under-resourcing and diverging national priorities that have characterized cooperation between data protection authorities.
    This may still leave concerned data protection authorities without the ability to meaningfully participate in the cooperation procedure and produce more careful and reasoned decisions.
    Under-enforcement of data protection law is especially troubling in connection to the ever-increasing scale of processing of personal data and the speed of technological development, which require strong regulatory oversight to strike a balance between the protection of fundamental rights and the free movement of data.
    The success of this reform will depend on whether data protection authorities will be willing and able to respect the strict procedural requirements provided and fulfill their constitutional duty, under EU primary law, to protect the fundamental right to data protection.
\end{abstract}