\begin{appendices}
\renewcommand{\thesection}{\Roman{section}}
\renewcommand{\thesubsection}{\thesection.\Alph{subsection}}
\section{AI Logbook}

\subsection{Use of Generative AI}

Generative AI will be used for minor tasks such as brainstorming, proofreading, and helping with surfacing relevant points in the literature to explore. 
Anthropic's Claude Sonnet and Opus models will be employed, with their thinking mode enabled. 
Claude will be used with a Pro subscription inside of Claude Code\autocite{HowClaudeCode} and Cowork.\autocite{GetStartedCowork}

Claude Code and Cowork allow the user to position the program, running locally, within a defined folder.\autocite{HowClaudeCode}
These tools allow for the model to have specific instructions on how to operate, giving it context into the precise work of the user that's being conducted within that folder.\autocite{HowClaudeRemembers}
This is done through a CLAUDE.md file.\autocite{HowClaudeRemembers}
These instructions constrain the model into acting in specific ways, in light of the guidelines for Responsible AI use, making Claude act as a brainstorm companion.
The CLAUDE.md file used during the writing process will be attached in the AI logbook.

Claude will also be able to read sources directly from Zotero through a \gls{mcp} server.\autocite{yu54yyyuZoteromcp2026}
An \gls{mcp} server is an open-source standard that allows AI tools to integrate databases and other tools in a structured way.\autocite{WhatModelContext}
Claude will be able to access notes and annotations done within Zotero, as well as the full text of the sources saved in the library.
This will help ground the model into the correct sources and limit hallucinations.

\end{appendices}