\begin{lscapepage}
\footnotesize\renewcommand{\arraystretch}{1.25}
\setlength\LTleft{0pt}\setlength\LTright{0pt}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\begin{longtable}{L{\dimexpr0.13\linewidth-12pt\relax} L{\dimexpr0.27\linewidth-12pt\relax} L{\dimexpr0.46\linewidth-12pt\relax} L{\dimexpr0.14\linewidth-12pt\relax}}
\hline \textbf{Date} & \textbf{Session focus} & \textbf{Claude's contribution} & \textbf{Tools} \\ \hline \endfirsthead
\hline \textbf{Date} & \textbf{Session focus} & \textbf{Claude's contribution} & \textbf{Tools} \\ \hline \endhead
\hline \multicolumn{4}{r}{\footnotesize\itshape continued on next page}\\ \endfoot
\hline \endlastfoot
2026-05-28 09:10–09:46 & Word count measurement and diagnostic script creation for thesis monitoring. & Measurement and tool-building (diagnostic script). & Bash, Read, Write \\ \hline
2026-05-28 14:17–14:18 & Diagnosed LaTeX underfull hbox warning in footnote typesetting. & Technical LaTeX diagnosis only; & Read, Bash \\ \hline
2026-05-28 18:43–22:07 & Chapter 4 and 5 structural planning to answer SQ3 and SQ4 (duty of care and duty to give reasons). & Structural diagnosis — identified gaps, mapped inputs, proposed framework. Created Obsidian planning document. & Read, Skill, Bash, AskUserQuestion, Write \\ \hline
2026-05-29 11:41–11:45 & Verify word count; identify and fix \textbackslash{}todo inflation in texcount. & Diagnosed bug in word-count infrastructure; edited the skill script (thesis\_stats.py). & Bash, Edit, Read \\ \hline
2026-05-29 19:08–19:50 & Restructuring chapter 4 (duty of care) after reading Obsidian notes, handwritten paper notes, and the current draft. & Structural diagnosis and source-finding (Zotero + local PDFs); mapped secondary literature to specific chapter sections; pinpointed case law passages for duty-of-care and record-vs-scope thesis; provided extensive pointers and hard questions for Matteo to write from. & Read, Bash, Zotero, Skill, ToolSearch, AskUserQuestion, Write \\ \hline
2026-05-30 19:55–21:19 & Structural review of chapter 4 against the planned outline written the day before. & Structural pointers, diagnostics of gaps against the outline, priority-ordered task list. & Bash, Read, Skill, AskUserQuestion \\ \hline
2026-05-31 10:38–2026-06-01 08:13 & Assessing chapter 4 structural integrity against plan; measuring word count; checking Obsidian structure note against current draft. & Structural review (plan vs. draft), flow diagnosis, word-count measurement. Provided pointers to gaps and diagnostic feedback. & Bash, Read, Skill \\ \hline
2026-06-01 09:27–09:54 & Word count check; investigate missing footnotes in OSCOLA LaTeX-to-docx converter output. & Deep debugging of OOXML infrastructure; edited converter repo script (/Development/oscola-tex-to-docx/sync-footnote-separators.py). & Bash, Read, Edit, Skill, AskUserQuestion \\ \hline
2026-06-01 11:07–14:31 & Word count analysis and debugging footnote-measurement script bug. & Measurement, diagnostics, and bug identification. & Bash, Read, Edit \\ \hline
2026-06-01 15:30–19:53 & Unblocking chapter 5 (duty to give reasons / SQ4) by clarifying organizing principle and reading AG Bobek's "empty shell" concern. & Structural pointers (Craig functions mapping); source-finding (Bobek, Bastos, C-97/23 P); identified organizing principle (what is reviewable in composite procedure?); drafted outline with explicit asks. & Bash, Read, Zotero, Skill, ToolSearch \\ \hline
2026-06-01 20:58 to 2026-06-02 15:13 & Building a thesis progress-visualization tool from git history using R and LaTeX. & Technical design and implementation. & Bash, Write, Read, Edit, Skill, AskUserQuestion, SendUserFile \\ \hline
2026-06-04 16:07–16:17 & Adding cross-reference labels to all chapter headings for internal hyperlinking. & Structural LaTeX mechanics only (label insertion, build verification). & Bash, Read, Skill, AskUserQuestion, Edit \\ \hline
2026-06-05 09:11–09:18 & Generating thesis progress report using git commit history and PDF building. & Progress report generation (infrastructure), commit message drafting. & Bash, SendUserFile \\ \hline
2026-06-05 11:19–11:37 & Matteo reported a LaTeX build warning (xdvipdfmx: Object @page.i already defined) appearing a couple of days ago. & LaTeX diagnostics (preamble mechanics, hyperref behavior), PDF debugging via xdvipdfmx output inspection, and applied one-line fix. & Bash, Read, Edit, AskUserQuestion \\ \hline
2026-06-05 11:37–11:44 & Word-count reporting and progress tracking at the start of chapter-2-rewrite branch. & Word-count statistics, progress reporting, and commit message drafting. & Bash, SendUserFile \\ \hline
2026-06-05 14:54–14:56 & Identifying and mapping redundancy in thesis introduction. & Structural analysis and pointer identification. & Read, Skill \\ \hline
2026-06-06 12:19–12:35 & Merged a single main.tex change from the chapter-2-rewrite branch onto main. & Git workflow and commit-message help only. & Bash \\ \hline
2026-06-08 09:43–09:48 & Word count measurement and thesis progress report generation. & Measurement, report generation, and infrastructure delivery & Bash, SendUserFile \\ \hline
2026-06-09 09:49–10:16 & Supervisor gave feedback on 1 June draft; Matteo wanted a breakdown of what to improve. & Feedback mapping, structural diagnosis. Offered source-finding (Zotero). & Bash, Read, Skill, Write \\ \hline
2026-06-09 10:31–13:41 & Analyzing supervisor feedback on draft thesis; creating improvement sheet for Obsidian. & Supervisor feedback analysis, structural diagnosis, scaffold creation for improvements. & Bash, Read, Edit, Write, Skill \\ \hline
2026-06-09 14:18–14:23 & Diagnosed VS Code memory usage after disabling GitHub Copilot. & System diagnostics and process analysis only; no code/thesis changes. & Bash \\ \hline
2026-06-10 10:08–12:53 & Processing supervisor feedback, planning thesis refocus on "peer pressure" framing, discovering and integrating sources. & Source discovery and research (Zotero, web search, PDF extraction); structural planning and framework development. Identified material and suggested organization & Bash, Zotero, Read, Edit, ToolSearch, WebFetch, Skill, Write \\ \hline
2026-06-11 14:37–14:49 & Fixing Zotero semantic search database setup — the MCP server was configured to use a broken Gemini API instead of the local embedding model. & Zotero infrastructure diagnosis and tooling fix. & Bash, Zotero, ToolSearch, Skill, Read \\ \hline
2026-06-11 14:50–14:53 & Source-finding for "peer pressure" concept (for chapter 1 "Peer Pressuring the LSA into Action?" section). & Source-finding (Zotero semantic search + academic databases). & Academic, ToolSearch, Zotero, Skill, Bash \\ \hline
2026-06-11 14:53–15:09 & Finding peer-pressure sources in Zotero and scaffolding new sections for chapter 1 and chapter 3. & Source-finding, structural scaffolding. & Bash, Zotero, ToolSearch, Skill, Read \\ \hline
2026-06-15 09:42–09:50 & Restructuring chapter 3 to integrate literature and cases; reading enforcement-flaws literature. & Source discovery, literature reading (Zotero search and fulltext retrieval); structural skeleton provided for section 1. & Read, Bash, ToolSearch, Zotero, Skill \\ \hline
2026-06-16 09:26–09:43 & Debugging LaTeX build errors (bibliography and reference issues). & LaTeX error diagnosis; explained the multi-pass rebuild process and why certain warnings are expected. & Bash, Read, Write, Edit \\ \hline
2026-06-18 11:20–12:08 & Word count on chapter-3-4-5-refocus branch; merging chapter-2-feedback branch; commit message for chapter 3 integration. & Word-count measurement; git/merge strategy advice and conflict resolution (did not execute edit but showed path); commit-message help. & Bash \\ \hline
2026-06-19 06:49–14:51 & Bibliography refactoring — moving from broken type= filters to biblatex categories to avoid keyword-field collisions with OSCOLA style requirements. & Bibliography mechanics (biblatex/oscola), entry categorization, mechanical .bib editing (merging duplicate keywords, finding outliers), file generation; Matteo approved schema and directions. & Bash, Read, Write, Edit \\ \hline
2026-06-19 15:02–15:09 & Matteo had staged a file reorganization (main.tex$\rightarrow$thesis.tex, bibliography folder creation) and needed reference updates. & LaTeX/project structure mechanics (reference locating, path updating, documentation consistency). & Edit, Bash, Read \\ \hline
2026-06-19 16:55–16:57 & Word count check; updated stats script to find renamed PDF (thesis.pdf instead of main.pdf). & Word-count measurement with tooling adaptation. & Bash, Read, Edit \\ \hline
2026-06-19 17:05–17:11 & Branch cleanup after merging chapter-3-4-5-refocus into main. & Git hygiene — branch diagnostics, cleanup execution. & Bash, AskUserQuestion \\ \hline
2026-06-20 09:39–2026-06-21 14:05 (multi-day) & Comprehensive chapter-3 restructuring; source-finding on cooperation file and judicial review; peer-pressure frame refinement. & Comprehensive progress audit, structural diagnosis for chapter 3, source-finding, argument mapping for peer-pressure lens and judicial review asymmetry. & Read, Bash, Zotero, Skill, Edit, ToolSearch \\ \hline
2026-06-21 13:35 & Citation key audit — counting unique sources used throughout the thesis. & Citation key extraction and counting. & Bash \\ \hline
2026-06-21 15:55–23:33 & Comprehensive chapter 4 source audit and revision checklist production. & Source audit, structural diagnosis, checklist production, progress tracking. Provided hypothesis questions, and mapped gaps. & Bash, Read, Edit, Write, Zotero tools, Skill, ToolSearch, SendUserFile, AskUserQuestion \\ \hline
2026-06-22 13:38–14:52 & Commit-message drafting, thesis progress reporting, parallel spellcheck, and acronym conversion (RRO). & Commit help, spellcheck reporting (via agents, no editing), LaTeX mechanics for acronym definition, word-count analysis. & Bash, Read, Agent, SendUserFile, Skill, Edit \\ \hline
2026-06-22 14:54–16:38 & Comprehensive chapter 5 structural refocus and deep source research (six sources read in full). & Structural diagnostic, source research (Zotero search), checklist generation. Claude provided diagnostic pointers, and task organization; Matteo provided the structure. & Bash, Read, Zotero, ToolSearch, Skill, Edit \\ \hline
2026-06-24 09:01–09:59 & Final word count measurement before thesis submission deadline (2026-06-26). & Measurement and statistics only. & Bash, Write, SendUserFile, Read, Skill \\ \hline
2026-06-24 10:00–19:14 & Matteo wanted a detailed gap analysis of Chapter 5 against his revision plan, with focus on the enforcement-system (not parties) layer. & Detailed structural analysis, gap mapping against revision plan, iterative feedback on each draft. Provided diagnostic tables and skeleton suggestions. & Read, Bash, Agent, Skill \\ \hline
2026-06-24 20:28–20:49 & Open todos audit and final word count check (2 days before 26 June submission deadline). & Todo extraction with context, word-count measurement. & Bash, Skill \\ \hline
2026-06-26 08:34–08:38 & LaTeX formatting for appendix: changing section numbering from letters to roman numerals. & LaTeX/bibliography mechanics (appendix package + counter renewal); explained side effects (ToC auto-update). Made two targeted edits to .tex file based on user request. & Edit, Read, Skill, AskUserQuestion \\ \hline
2026-06-26 11:37–11:40 & Matteo wanted to remove the todonotes package from the LaTeX preamble. & Package-dependency troubleshooting, LaTeX build diagnostics. Identified and fixed a latent transitive dependency bug. & Bash, Edit, Read \\ \hline
\end{longtable}
\end{lscapepage}