\section[The General Principle of Good Administration as a Constitutional Safeguard]{The General Principle of Good Administration\\as a Constitutional Safeguard}

General principles of \gls{eu} law represent `general, comprehensive [...], and naturally inherent'\autocite[42]{audiolux} constitutional principles that permeate the Union legal order.\autocite[7]{tridimasGeneralPrinciplesEU2020} 
They appeared as a natural consequence of the political and legal identity of the \gls{eu},\autocite[2]{TEU} a nascent legal order created by international treaties that provide a framework for incremental integration among the \glspl{ms}.\autocite[17--18]{tridimasGeneralPrinciplesBook2006}

The 1957 Treaty of Rome already empowered the Court of Justice to review legal acts,\autocite[173]{treatyofRome} but the Treaty did not provide detailed guidance on the standard of review to be applied by the court.\autocite[265--266]{craigFoundations2018} 
Therefore, in order to fill in the gaps, and in the interest of ensuring that Union institutions would face scrutiny,\autocite[269]{craigFoundations2018} the Court of Justice developed a system of background principles of law inferred from the legal and constitutional traditions of the \glspl{ms}.\autocite[269--270]{craigFoundations2018}
The work of the Court in this regard was not a mere comparative study of the principles that underpinned \glspl{ms}' legal systems, but rather an investigation of the `best' interpretation of shared values in light of the objectives of the Treaties.\autocites[270]{craigFoundations2018}[See also][point 69]{audioluxAG}

General principles of \gls{eu} law concern both principles derived from the concept of the rule of law as understood in a liberal democracy,\autocites[9]{tridimasGeneralPrinciplesEU2020}[point 69]{audioluxAG} and principles derived from the institutional design of the \gls{eu}.\autocite[11--12]{tridimasGeneralPrinciplesEU2020}
This chapter will focus on the general principle of good administration, its history, and its role as a constitutional safeguard in the European composite administrative enforcement model.

\newpage
\subsection{The Concept of 'Good Administration'}

`Good administration' appears in the Union legal order both as a general principle of \gls{eu} law and as a fundamental right enshrined in the \gls{cfr}. 
This section will deal with the two sources of this right in the \gls{eu}'s legal system, and provide an analysis between the two.

\subsubsection{Good Administration as a General Principle of EU Law}

One of the first instances where the Court of Justice undertook the task of `filling in the gaps' left by the Treaties was in \emph{Algera}, one of the earliest cases of the Court, dealing precisely with administrative matters:

\begin{quote}
    The possibility of withdrawing such measures is a problem of administrative law, which is familiar in the case-law and learned writing of all the countries of the Community, but for the solution of which the Treaty does not contain any rules.
    Unless the court is to deny justice it is therefore obliged to solve the problem by reference to the rules acknowledged by the legislation, the learned writing and the case law of the member countries.\autocite[para A III]{algera}
\end{quote}

Throughout the decades, the principle of good administration has been greatly developed by the \gls{cjeu}.
The principle appears in the case law in various forms: as `good',\autocite[para 1A]{alvis} `sound',\autocite[para A II]{societesarre} `proper'\autocite[paras 7 and 12]{burban} administration. 
While today the concept of good administration is widely recognized in the jurisprudence to constitute a general principle of Union law, its early years were marked by a reluctance to recognize it as such.\autocite[66--68]{mihaescu-evansRightGoodAdministration2015} 

The resistance may have stemmed from the fact that the concept of good administration was referred to in inconsistent ways by the judiciary. 
For example, good administration appeared in the case law as an `obligation',\autocite[paras 245 and 279]{volkswagenAg} a `duty',\autocite[126]{bundesverbandDeutscherBanken} a `rule',\autocite[4]{vanEick} a `requirement',\autocite[71]{vba} or a `principle'.\autocite[24--31]{tetraPak} 
A narrow interpretation of the principle is also observed in cases where the Court of Justice held that while violations of the principle of good administration might have occurred, they did not have an effect on the final legality of the act in question.\autocite[67--68]{mihaescu-evansRightGoodAdministration2015}

Nowadays, the principle of good administration is widely recognized as a general principle of Union law.\autocite[210]{hotelCipriani} 
However, the jurisprudence and the literature concede that good administration is an `umbrella' concept, without a precise delineation as a standalone principle,\autocite[points 60--61]{commVScottAG} and that it is defined by the more precise procedural rights and requirements it encompasses.\autocite[73--74]{mihaescu-evansRightGoodAdministration2015}

To discern the content of the general principle of good administration, we turn to Article 41 \gls{cfr}, which provides a number of sub-principles that can be understood as falling under the umbrella of good administration.
\autocites[75--76]{mihaescu-evansRightGoodAdministration2015}[127]{Tillack}
\todo[inline]{Supervisor: I would say that these sub-principles were already determined before the entry into force of the CFR, so I would perhaps rephrase. Especially as Art. 41 CFR is not applicable to the MSs, which you rightly point out later on in the chapter. Perhaps you should say that earlier on already.}

\subsubsection{The Charter of Fundamental Rights and Article 41 CFR}

The Charter of Fundamental Rights was first solemnly proclaimed in 2000 as a political document,\autocite[27--28]{fraHandbookDataprotection2018} and given binding force in 2009 with the entry into force of the Lisbon Treaty.\autocite[6(1)]{TEU}
Section V, on citizens' rights, contains Article 41, titled `Right to good administration'.\autocite[41]{CFR}
Article 41 \gls{cfr} is directly connected to the general principle of good administration established in the case law, and based on the `existence of the Union as subject to the rule of law'.\autocite[Explanation on Article 41]{ExplanationsCFR}

Article 41 \gls{cfr} provides a number of rights inserted under the umbrella of `good administration'.
As the \gls{cjeu} outlined in \emph{Tillack}:

\begin{quote}
    [The principle of sound administration] constitutes the expression of specific rights such as the right to have affairs handled impartially, fairly and within a reasonable time, the right to be heard, the right to have access to files, or the obligation to give reasons for decisions, for the purposes of Article 41 of the Charter of fundamental rights of the European Union\autocite[127]{Tillack}
\end{quote}

The wording of the Court hints at a non-exhaustive list of principles,\autocite[91--92]{hofmannEvansRelationCharterFundamental2013} similarly to Article 41(2) \gls{cfr}, which states that the right `includes' a number of specific procedural rights.\autocite[41(2)]{CFR}

One important note to make is that Article 41(1) \gls{cfr} explicitly limits the institutional scope of the Charter right to good administration.
The article circumscribes the right to good administration to the `institutions, bodies, offices and agencies of the Union'.\autocite[41(1)]{CFR}
The Charter right also presents a narrower material scope compared to the case law principle, making references to the `individual', the right to access `his or her' file, or an `individual measure'.\autocite[86--87]{hofmannEvansRelationCharterFundamental2013}

Therefore, attention must be paid to the way these two sources of the right to good administration interact with each other.

\subsubsection[\texorpdfstring{The Coexistence of Good Administration as a General Principle\\and as a CFR Right}{The Coexistence of Good Administration as a General Principle and as a CFR Right}]{The Coexistence of Good Administration as a General Principle and as a CFR Right}

The relationship between good administration as a general principle of Union law and as a Charter right has been the subject of discussion both in the literature and in the case law.\autocite{mihaescu-evansRightGoodAdministration2015}

As highlighted above, the right to good administration contained in Article 41 \gls{cfr} is narrower in scope than the general principle of good administration as developed in case law.\autocite[148]{mihaescu-evansRightGoodAdministration2015}
The limitations in the personal, material and institutional scope of the Charter right are problematic especially because, as it is the case with data protection law, \glspl{ms} are most often the enforcers of Union law.\autocite[150]{mihaescu-evansRightGoodAdministration2015}
\todo[inline]{Supervisor: Why problematic? If the principles are generally broader in scope, isn't it a good thing that those are applicable to the MSs? They have also equal weight, right?}

The \gls{cjeu} has however clarified that while the right contained in the Charter does not apply to \glspl{ms},\autocite[68--69]{ysVMinister} when invoked as a general principle, good administration must be observed by \glspl{ms} when acting in the scope of Union law.\autocite[50]{HNvMinisterJustice} 

\subsection[Good Administration in the European Composite Administrative Enforcement Model]{Good Administration in the European Composite Administrative\\Enforcement Model}

As explored earlier, the \gls{eu} presents a unique institutional design, where the Union legal order supplements the national legal orders of the \glspl{ms}. 
Consequently, \glspl{ms} are often the main enforcers of Union law, which is supposed to be applied uniformly and predictably across the territory of the Union.
In the realm of administrative enforcement, the constitutional structure of the \gls{eu} generates a `multi-level', or `composite' administrative procedure, where the administrations of different \glspl{ms} and Union institutions might be collectively responsible for carrying out enforcement on an individual case.\autocite[91--92]{mihaescu-evansRightGoodAdministration2015}
\todo[inline]{Supervisor: I would include a reference here to Hofmann.}

This section will highlight the features of the `integrated' administrative model of the \gls{eu}, and the role of the principle of good administration in ensuring the rule of law across the Union legal order.
\todo[inline]{Supervisor: It would be good to take the reader a bit by the hand in the first part of this chapter. Here it would be good, for example, to link back to your research topic and briefly mention why this is important to understand.}

\subsubsection{The Composite Administrative Enforcement Model}

\gls{eu} law presents many policy areas where enforcement is shared among different institutional actors at the Union and national level.\autocite[138]{hofmannCompositeDecisionMaking2009}
We can distinguish between two types of cooperation: horizontal, between actors on the same level of government (such as National Competition Authorities of different \glspl{ms}), and vertical, between actors on different levels of government (such as a National Competition Authority and the European Commission).
\todo[inline]{Supervisor: Perhaps link to the GDPR case? Horizontal between DPAs and vertical between DPAs and the EDPB?}

The `composite' nature of administrative procedure in the \gls{eu} is a natural by-product of a decentralized European administration governed by subsidiarity and transjurisdictional cooperation.\autocite[91--92]{mihaescu-evansRightGoodAdministration2015}
However, issues arise from the convergence of different administrative procedural requirements established by the different legal frameworks the actors are subject to, with effects on individual cases and the overall effectiveness of judicial oversight of administrative action.\autocite[137]{hofmannCompositeDecisionMaking2009} 
Hofmann observes how the lack of systematic harmonization of administrative law exacerbates this problem in the \gls{eu}.\autocite[137]{hofmannCompositeDecisionMaking2009}

The principle of good administration however could function as a tool to fill in the gaps left by the lack of comprehensive harmonization, acting as a tool to ensure the respect of the rule of law across the complex administrative procedures taking place in the \gls{eu}.\autocites[92--93]{mihaescu-evansRightGoodAdministration2015}[351]{nehlGoodAdminProceduralRight2009}

\subsubsection{The `Judicialization' of the Administration}
\todo[inline]{Supervisor: If you get difficulties with word count, I would say that this could also be integrated shorter in section 2.2.1.}

The right to good administration is explicitly linked to the right to effective judicial protection established by Article 47 \gls{cfr}.\autocites[Article 41]{ExplanationsCFR}[79--80]{mihaescu-evansRightGoodAdministration2015}
The principles of good administration mirror the principles of effective judicial protection, which is also viewed as an `umbrella' concept.\autocite[79--80]{mihaescu-evansRightGoodAdministration2015}
Establishing strong procedural safeguards at the administrative stage is crucial to ensure the rule of law is upheld during the administrative procedure, and to enable an appellant to successfully challenge the actions of the administration at the judicial stage.\autocite[79--80]{mihaescu-evansRightGoodAdministration2015}
For example, an appellant's right to effective judicial protection would be undermined if a judge would not be able to ascertain the logic and reasoning behind a decision of the administration, hence the need for the administration to provide reasoned decisions.\autocite[81]{mihaescu-evansRightGoodAdministration2015}

The importance of upholding the principle of good administration as a feature of the administrative system is also augmented by the increasing complexity of the European administrative model.\autocite[86--87]{mihaescu-evansRightGoodAdministration2015}
While it had been accepted by the Courts that the right to effective judicial protection would not be infringed upon as long as an ex-post judicial review of administrative acts was available, the institutional complexity of the \gls{eu} has made it increasingly difficult to challenge the discretionary power of national administrations.\autocite[85--86]{mihaescu-evansRightGoodAdministration2015}
In fact, it would be impossible for the \gls{eu} judiciary to substitute itself to national administrative authorities and replace their power of appraisal, given the delicate institutional balance that exists between Union institutions and the \glspl{ms}.\autocite[87]{mihaescu-evansRightGoodAdministration2015}

However, in a composite administrative system, with national authorities whose decisions are sometimes liable for the protection of the fundamental rights of over 450 million people, integrating good administration as a design of administrative processes may be a counterweight to the discretionary powers of administrations, leading to better outcomes for the rule of law and the protection of fundamental rights.

\subsubsection{A Limitation on the Administration's Discretionary Power}
\todo[inline]{Supervisor: Isn't this a bit repetitive?}

Good administration concerns both the procedural rights of the individual when facing the administration, as well as the effectiveness of the administrative system concerned.\autocite[88]{mihaescu-evansRightGoodAdministration2015} 
These are, respectively, the subjective and objective dimensions of the principle of good administration.

The two dimensions of good administration, according to Mihaescu-Evans, should not be understood as mutually exclusive, but two sides of the same coin.\autocite[114]{mihaescu-evansRightGoodAdministration2015}
In fact, the appropriate protection of each individual's procedural rights would be able to ensure an administration that is effective and that works within the bounds of the rule of law, leading to better enforcement outcomes for the integrity of the legal system and the protection of individuals.\autocite[116--117]{mihaescu-evansRightGoodAdministration2015}

\subsection{The Administration's Duty to Deliver Careful and Reasoned Decisions}

This section will explore the requirements of the sub-principles of good administration that are most relevant to the enforcement model of the \gls{gdpr} as amended by the Procedural Regulation, namely the duty of care and the duty to give reasons.

\subsubsection{The Duty of Care of the Administration}
\todo[inline]{Supervisor: Sections 2.3.1 and 2.3.2 are crucial components of your thesis, as they follow the benchmarks of your analysis. Expanding a bit if possible --- i.e., is there more relevant literature and case law --- would be very helpful.}

The duty of care of the administration is a key aspect of the principle of good administration.
The principle of good administration was initially developed by the \gls{cjeu} when referring to the duty of the administration to be careful in its information-gathering prior to reaching a decision.\autocite[195]{hofmannAdministrativeLawPolicy2012}
\todo[inline]{Supervisor: And decision-making, see comment above (broader definition of duty of care).}
Today, the duty of care is a requirement on the administration to impartially and carefully establish the facts and points of law relevant to a case when acting in its adminstrative capacity to deliver a decision.\autocites[point 15]{NolleAG}[195]{hofmannAdministrativeLawPolicy2012}

The duty of care is especially relevant in cases where the administration enjoys a vast power of appraisal,\autocite[14]{TUM} acting as a check to the discretionary power by providing a minimum standard of conduct in decision-making and reason-giving.\autocite[216]{hofmannDemkovaPrinciplesProcJustice}
The administration must conduct itself in a manner that `serious',\autocite[59]{Metropole}, careful, cautious,\autocite[93]{Masdar} in making its decisions.
When the administration fails to act with care, it may be found to have been `negligent', and liable for its lack of care when harm results from it.\autocite[90--91]{Masdar}

The duty of care applies to all stages of an administrative proceding, where for example a multi-step procedure exists where a preliminary investigation may be followed up by a more in-depth one.\autocite[447]{hofmannAdministrativeLawPolicy2012} 
The administration must be careful in, for example, whether to open an investigation after a complaint was filed, with the General Court stating that `the diligent and impartial treatment of a complaint is associated with the right to sound administration which is one of the general principles that are observed in a State governed by the rule of law'.\autocite[48]{maxMobil}

According to multiple scholars, the duty of care includes the obligation of the administration to take decisions based on all relevant facts, in a way that ensures data is accurate, and following a thorough examination of the case.\autocites[412]{tridimasGeneralPrinciplesBook2006}[35]{mustertCrossborderEnforcementGDPR}[84]{hofmannEvansRelationCharterFundamental2013}
The exercise of the duty of care must lead the administration to possess the `most complete and reliable information' for making a decision.\autocite[90]{Scott}
The duty of care is therefore connected to the ability of the administration to collect, manage, and assess the information necessary to reach a decision.\autocite[36]{mustertCrossborderEnforcementGDPR} 

The scope of information-gathering, Hofmann argues, includes `all relevant aspects of a case which may have a bearing on the adoption of a measure including in cases where the decision maker has discretion.'\autocite[98]{hofmannDutyCareEU2020}
Importantly, the administration cannot simply rely on the information provided to it by the parties to the procedure, but must use its investigative powers to examine facts not initially brought by the complainant.\autocites[62]{Sytraval}[36]{mustertCrossborderEnforcementGDPR}

In synthesis, the principle of good administration serves the judiciary as a tool of to assess the legality of a decision by examining the quantity and quality of the information at the disposal of the administration, and the way it was used to reach a decision.\autocite[400]{mihaescu-evansRightGoodAdministration2015}
The completeness of the information at the disposal of the administration is crucial to enable it to provide reasoned decisions, another key aspect of the principle of good administration.

\subsubsection{The Duty to Give Reasons of the Administration}
\todo[inline]{Supervisor: This is very much about the rationale of reason-giving. But what is sufficient reasoning? This framework should be developed here.}

The duty to give reasons refers to the obligation for authorities to communicate the reasons motivating their decisions to the parties affected.\autocite[38]{mustertCrossborderEnforcementGDPR} 
As discussed by Craig, the duty to give reasons has three different functions.

First, from the viewpoint of the citizen, the right to receive reasons for a decision impacting them is important to make them understand why a decision has been adopted.\autocite[369]{craigProcess2018}
Reason-giving makes the administration transparent and accountable to the citizen, and allows them to ascertain whether the decision can be challenged in the courts.\autocite[38]{mustertCrossborderEnforcementGDPR}

Second, from the viewpoint of the judiciary, the duty of the administration to give reasons is crucial to enable judicial review of administrative acts.\autocite[370]{craigProcess2018}
As discussed earlier, the right to effective judicial protection is closely linked to the information the judiciary receives about the grounds for an administrative decision.

Finally, from the viewpoint of the administration itself, the exercise of duty to give reasons is important as a check on the quality of the assessments that led to the decision.\autocites[38]{mustertCrossborderEnforcementGDPR}[370]{craigProcess2018}
This requirement, as argued by Opdebeek and De Somer, `forces administrative bodies to reflect on the legality, quality, rationality, reasonableness and fairness of their decisions, which may prevent bad decision-making in the first place.'\autocite[106]{opdebeekDutyGiveReasons2016}

Therefore, it follows that a diligent investigation of the facts and the points of law by the administration is necessary to ensure that the administration can provide reasoned decisions. 
In turn, the obligation to provide reasoned decisions forces the administration to be careful about the logic and the facts that contribute to their assessments, providing a check before decisions are issued.

\subsection{The Collective Duty of Data Protection Authorities to Deliver Careful and Reasoned Decisions}
\todo[inline]{Supervisor: I doubt a little bit if this section is logically placed here. Perhaps integrate in the text above or discuss in the next chapter. Let's discuss tomorrow.}

After considering the aspects of the duty of care and the duty to give reasons within the principle of good administration, we can now turn to the specific features of the \gls{gdpr}'s enforcement model.
This section will explore the design of the cooperation and consistency mechanism of the \gls{gdpr} as a decentralized enforcement network, and explore the role of the duty to give careful and reasoned decisions in its context.

\subsubsection{The GDPR's Cooperation Mechanism as an Enforcement Network}

Within the cooperation and consistency mechanism of the \gls{gdpr}, we notice a network of horizontal cooperation between \glspl{dpa}, with the \gls{edpb} serving as a vertical norm-setter and decision-maker.\autocites[140]{hofmannCompositeDecisionMaking2009}[18--22]{mustertCrossborderEnforcementGDPR}
As Mastenbroek and Sindbjerg Martinsen argue, continuous interaction within an enforcement network may foster knowledge exchange, an increase in mutual trust, and persuasion among its members.\autocite[428]{mastenbroekFillingGapEuropean2018}

The cross-border enforcement model of the \gls{gdpr} is heavily based on cooperation through information exchange.\autocite[140]{hofmannCompositeDecisionMaking2009}
The \gls{lsa} and \glspl{sac} must reach consensus on every stage of the procedure,\autocite[465]{hofmannmustertScholtenHandbook2023} and in doing so, they must share information,\autocite[60(1)]{GDPR} provide mutual assistance to each other, and participate in joint operations.\autocite[60(2)]{GDPR}
On the other hand, the \gls{edpb} acts as a central node in the enforcement network, providing support to the \glspl{dpa} while discharging their duties.\autocite[467]{hofmannmustertScholtenHandbook2023} 
\glspl{dpa} can request opinions or binding decisions from the \gls{edpb}, and use the \gls{edpb} as a dispute settlement forum in case of disagreements during the cross-border enforcement procedure.\autocite[467--468]{hofmannmustertScholtenHandbook2023}

\subsubsection{The Collective Duty of Enforcement}

As explored in section 2.2.1, the design of the \gls{gdpr}'s cross-border enforcement model qualifies as a composite administrative procedure, where the \gls{lsa} leads the procedure and is the sole interlocutor of the accused, but has to consult \glspl{sac} to reach a common enforcement outcome.\autocite[63]{facebookBelgium}

As the \gls{cjeu} affirmed in \emph{Facebook Belgium}, the one-stop-shop mechanism of the \gls{gdpr} does not relieve the \glspl{dpa} of their collective duty to ensure the application of the \gls{gdpr} and the protection of data subjects' fundamental right to data protection.\autocite[63--68]{facebookBelgium} 
AG Bobek further argued that an \gls{lsa} must act as a \emph{primus inter pares}, acting only with the consent of the \glspl{sac}.\autocite[point 111]{fbBelgiumAgBobek}

\todo[inline]{Talk about subsidiarity and proximity here? + connect to Hofmann on composite decision making}

However, a collective duty of enforcement can only be discharged by all \glspl{dpa} if they are able to access all relevant information about a case, and if there is clear understanding behind the choices made in the information-gathering stages.

\subsubsection{Locating the Duty of Care and the Duty to Give Reasons in the GDPR's Enforcement Model}

The duty to deliver careful and reasoned decisions becomes particularly important in the integrated administrative model of the \gls{gdpr}, especially when considering the objective facet of the principle of good administration.\autocite[114]{mihaescu-evansRightGoodAdministration2015}

The duty of care of \glspl{dpa} has already appeared in the case law of the \gls{cjeu}. 
The Court has ruled that \glspl{dpa} must examine complaints `with all due diligence',\autocites[63]{SchremsI}[68]{SCHUFA} meaning that they must `examine the nature of that complaint as necessary'.\autocites[109]{SchremsII}[37]{mustertCrossborderEnforcementGDPR} 
Under the framework of the \gls{gdpr}, this duty of care will only be discharged through extensive information-gathering and cooperation between the \gls{lsa} and the \glspl{sac}, so that all relevant aspects of a case can be taken into account by the \gls{lsa} when drafting a decision.

Similarly, recalling the functions of the duty to give reasons laid out by Craig,\autocite[369--370]{craigProcess2018} a close collaboration between \glspl{dpa} in a cross-border case could aid the integrated system of the \gls{gdpr} as a whole to check the quality of its assessments, ensuring that the final decision is based on lawful and well-reasoned grounds.
The duty to give reasons also comes into play in the context of dispute settlement, where the \gls{edpb} needs to provide a binding decision based on the assessments and the objections of the \glspl{dpa} involved in the procedure.

However, as evidenced by Lynskey and Gentile, when an \gls{lsa} narrows the scope of the investigation and fails to collect evidence on a contested point of law, the \gls{edpb} might not have enough elements to conduct a thorough review of the case, the quality of the draft decision of the \gls{lsa}, and the reasoning behind the objections of the \glspl{sac}.\autocite[810]{gentileDeficientDesign2022}

The Procedural Regulation introduces a number of procedural steps to mandate consensus-building and information exchange between \glspl{dpa}, and will be dealt with in the next chapter.