\section[The Cooperation and Consistency Mechanism Pre-- and Post--Procedural Regulation]{The Cooperation and Consistency Mechanism\\Pre-- and Post--Procedural Regulation}

The cooperation and consistency mechanism forms the core of the enforcement design of the \gls{gdpr} for cross-border processing, through a network of horizontal cooperation between \glspl{dpa} with a vertical decision-maker and norm setter in the \gls{edpb}.
Networks of horizontal cooperation like the \gls{gdpr}'s, as seen in section 1.1.3, may increase the quality and speed of enforcement through a mechanism of peer pressure.\autocites[466]{hofmannmustertScholtenHandbook2023}[428]{mastenbroekFillingGapEuropean2018}
This chapter will first explore the pre--reform enforcement mechanism, with special regard to the ways in which \glspl{sa} have been able to influence one another throughout the process.
It will map the ways in which the cooperation and consistency mechanism has failed to deliver effective avenues to steer enforcement into more effective outcomes, as detailed in the literature and individual enforcement cases.
Finally, the chapter will analyze the key changes to the flow of cooperation introduced by the Procedural Regulation, with a view of understanding whether the reform will bring more (effective) legal avenues to push enforcement forward through peer pressure.

\subsection{The Tasks and Powers of Supervisory Authorities in Cross-Border Cases}

Under the \gls{gdpr}, \glspl{ms} are required to establish independent\autocite[52]{GDPR} \glspl{sa}, entrusted with monitoring the application of the Regulation.\autocite[51]{GDPR} 
Importantly, the \gls{gdpr} requires \glspl{sa} to all contribute to the application of the Regulation throughout the Union.\autocite[51(2)]{GDPR} 
The \gls{sa} is competent for enforcing the Regulation within the territory of its \gls{ms},\autocite[51(1)]{GDPR} with well-defined tasks\autocite[57]{GDPR} and powers\autocite[58]{GDPR}.
Among the tasks of the \gls{sa} figure: the monitoring of the application of the \gls{gdpr};\autocite[57(1)(a)]{GDPR} the handling of complaints;\autocite[57(1)(f)]{GDPR} the cooperation with other \glspl{sa} to ensure the consistency of application of the \gls{gdpr};\autocite[57(1)(g)]{GDPR} and the handling of investigations, including with the input of other \glspl{sa}.\autocite[57(1)(h)]{GDPR}
The powers of the \gls{sa} include investigative, corrective, advisory powers, as well as the power to engage in legal proceedings.\autocite[58]{GDPR}

\subsubsection{Competence Allocation in Cross-Border Enforcement Cases}

In the case of cross-border processing, the \gls{lsa} becomes competent to supervise the processing activities of the controller or processor when its main or single establishment is in its territory.\autocite[56(1)]{GDPR}
The \gls{lsa} becomes the sole interlocutor of the controller or processor responsible for the cross-border processing at examination.\autocite[56(6)]{GDPR}
Cross-border processing refers to either the processing of personal data by a controller or processor with multiple establishments across the Union, or the processing of personal data by one entity that `affects or is likely to affect' data subjects in more than one \gls{ms}.\autocite[4(23)]{GDPR}

Other \glspl{sa} then may become `concerned'.\autocite[4(22)]{GDPR} 
An \gls{sac} is concerned when either: the processing in question is carried out by an establishment in its territory; the data subjects residing in the territory of its competence are affected by the processing; or when a complaint has been lodged with it.\autocite[4(22)(a)--(c)]{GDPR}

\subsection{The Pre-Reform Cooperation and Consistency Mechanism}

\subsubsection{The Cooperation Procedure}

Articles 60 and 61 \gls{gdpr} set out the cooperation procedure between the \gls{lsa} and the \glspl{sac} in cases involving cross-border processing. 
Fundamentally, the \gls{lsa} is required to cooperate with the \glspl{sac} to reach consensus, with the authorities exchanging `all relevant information with each other'.\autocite[60(1)]{GDPR}
This endeavour does not form a legal obligation, but rather an objective, and reflects the wishes of the legislator to bring about a system mandatory cooperation between the \glspl{sa}.\autocite[39--43]{edpbGuidelines022022}

While investigating potential infringements, \glspl{sa} can make reciprocal requests for mutual assistance.\autocite[60(2), 61(1)]{GDPR} 
Mutual assistance includes requests for information, investigations, and inspections.\autocite[61(1)]{GDPR}
\glspl{lsa} can also make use of joint operations, where the \gls{lsa} and \glspl{sac} can carry out joint investigations, inspections, or monitor the application of corrective measures.\autocite[60(2), 62]{GDPR}
The ability to exchange information closely relates to forming consensus on a case, and it is a necessary obligation to respect to enable all \glspl{sa} to form an opinion on the case and fulfill their roles under the law.\autocite[44--45]{edpbGuidelines022022}

These tools have encountered practical hurdles as evidenced by Hofmann and Mustert.
While peer pressure through information exchange duties may push enforcement forward, \glspl{dpa} have experienced the formal tools offered by the pre-reform system as a burden on their limited resources.
Without information, however, \glspl{dpa} lack the knowledge necessary to be an active part of the decision-making process, thus limiting their ability to advocate for a certain enforcement outcome.

\paragraph{Relevant and Reasoned Objections}
A lack of sufficient information becomes especially problematic in the final stages of the cooperation procedure.
In fact, when the \gls{lsa} has terminated its investigation, it must submit a draft decision to the \glspl{sac} and consult their views on the draft.\autocite[60(3)]{GDPR}
\glspl{sac} have four weeks to submit a `relevant and reasoned objection' to the draft.\autocite[60(4)]{GDPR}
An objection is regarded as `relevant and reasoned' if `clearly demonstrates the significance of the risks posed by the draft decision' concerning the fundamental rights and freedoms of data subjects.\autocite[4(24)]{GDPR}
The \glspl{sac} have a high threshold to clear, with the objection presenting `clearly and precisely' the link between the consequences the draft decision will bring about, and the significance of the risk for data subjects' fundamental rights and freedoms.\autocites[18--19]{edpbGuidelines092020}[811]{gentileDeficientDesign2022}
General comments are not enough to meet this threshold,\autocite[417]{mustertEDPBsteppingup2021} thus demanding an in-depth engagement with each individual case on the part of the objecting \gls{sa}.\autocites[811]{gentileDeficientDesign2022}[466]{hofmannmustertScholtenHandbook2023}

An \gls{lsa} can follow the objections presented, and adjust the draft decision accordingly.\autocite[60(5)]{GDPR}
In reality, \glspl{sac} have struggled to meet the high threshold required.
In those cases, when the \gls{lsa} does not follow the objections, or it does not believe that they are relevant or reasoned, the matter must be referred to the consistency mechanism.\autocite[60(4)]{GDPR}

If the period of four weeks lapses, and no objections are raised, the \gls{lsa} can adopt the decision and the \glspl{sac} are deemed to have agreed to it.\autocite[60(6)]{GDPR} 
When a decision is adopted, or a complaint is dismissed or rejected, the \glspl{sac} who received the original complaint are tasked with notifying the complainants of the decision.\autocite[60(7)--(8)]{GDPR}
When a complaint is only followed in part, the \gls{lsa} must inform the complainant of the points raised where it chose to act, while the \gls{sac} must adopt a separate decision dismissing the rejected parts of the complaint.\autocite[60(9)]{GDPR}

\subsubsection{The Consistency Mechanism}

The consistency mechanism is an extension of the cooperation procedure, meant to bring a consistent enforcement of the \gls{gdpr} across the Union.\autocites[63]{GDPR}[68]{googlevCNIL}[52]{facebookBelgium}
It is laid out in Articles 63--67 \gls{gdpr}, and can be triggered in various ways in the context of cross-border enforcement.

Firstly, the \gls{edpb} can issue non-binding opinions on matters of general application on the request of an \gls{sa}, its own chair, or the Commission.\autocite[64(2)]{GDPR}
This forms a first tool at the disposal of a \gls{sac} in case an \gls{sa} refuses to comply with a request for mutual assistance or joint operations as discussed by AG Bobek in \emph{Facebook Belgium}.\autocites[point 120--121]{fbBelgiumAgBobek}[64(2)]{GDPR}
While an opinion is non-binding, a binding decision can be requested in the event that an opinion is not followed, creating an implicit --  albeit cumbersome to operationalize -- incentive for the authorities to follow that opinion.\autocites[points 120--122]{fbBelgiumAgBobek}[318]{matosbrandaoOnestopshopEuropeanData2023}

In addition to issuing binding decisions on the failure to follow an \gls{edpb} opinion, the Board can issue binding decisions in a number of instances.\autocite[65(1)]{GDPR}
The \gls{edpb} can resolve disputes between \glspl{sa} when there are conflicting views as to which \gls{sac} is competent for the main establishment of a controller or processor.\autocite[65(1)(b)]{GDPR}
Most importantly, the \gls{edpb} functions as a dispute settler in the cooperation procedure, as mentioned in the previous section.
When a \gls{sac} raises a relevant and reasoned objection to a draft decision of the \gls{lsa}, and the \gls{lsa} does not follow the objection, the \gls{lsa} is obliged to refer the matter to the \gls{edpb}.\autocite[65(1)(a)]{GDPR}

It is important to note that the \gls{edpb} does not have investigative powers of its own, and functions as a supporter of the main enforcers of the \gls{gdpr}, which remain the national \glspl{dpa}.\autocite[491--492]{britobastosCentralisedGeneralData2023}
The \gls{edpb} is therefore bound by the information gathered by the national \glspl{sa},\autocite[468]{hofmannmustertScholtenHandbook2023} and can only rule on the matters that national authorities have asked it to rule upon.\autocite[492]{britobastosCentralisedGeneralData2023}

The limitations of the \gls{edpb}'s role have already emerged in practice where the Board was not able, in a number of cases, to side with the \glspl{sac} who successfully brought a reasoned and relevant objection.
Examples include the \emph{Twitter Ireland} case, where the Board recognized that while further infringements of the \gls{gdpr} might have been committed by Twitter, the limited scope of the investigation started by the Irish \gls{sa} meant that it did not have elements to rule upon such alleged infringements.\autocite[132]{edpbTwitter}
In a similar vein, the \gls{edpb} compelled the Irish \gls{sa} to initiate a new investigation in the case of \emph{Instagram Ireland}, because it found that too little evidence was available to rule on potential infingements of Article 9, which were excluded from the scope of the investigation.\autocite[200--203]{edpbInstagram}
In \emph{Instagram}, the Board particularly emphasized that a reduction in the scope of the investigation constrains the capacity of the \glspl{sac} to effectively act and tackle the risks to data subjects, and impairs the \gls{edpb}'s capacity to ensure a consistent application of the \gls{gdpr}.\autocite[200]{edpbInstagram}

Conversely, in the case of \emph{WhatsApp Ireland}, the Board was able to accept the relevant and reasoned objection of the Italian \gls{sa}.\autocite[183--201]{edpbWhatsApp}
However, the Board was only able to expand the legal scope of the decision insofar as the investigated breaches of the \gls{gdpr} concerning informational duties by WhatsApp were systematic and grave enough to rise to a violation of the fundamental principle of data processing in Article 5(1)(a).\autocite[198--201]{edpbWhatsApp}
All three decisions highlight the importance of early scope setting and consensus, in order to effectively enable the \glspl{sa} to meaningfully participate in the enforcement procedure, as discussed in the previous section.\autocites[132--136]{edpbTwitter}[225]{edpbWhatsApp}[200-203]{edpbInstagram}[468]{hofmannmustertScholtenHandbook2023}

Finally, another instrument of pressure at the disposal of \glspl{sac} is found in the urgency procedure, when an \gls{sa} believes that there is an urgent need to act to protect the fundamental rights and freedoms of data subjects.\autocite[63]{facebookBelgium}
The urgency procedure allows any \gls{sa} to adopt urgent provisional measures, with a maximum effect of three months, on the territory of their \gls{ms}.\autocite[66(1)]{GDPR}
\glspl{sac} must inform all the other \glspl{sa}, the \gls{edpb}, and the Commission of their urgent provisional measure, and may request an urgent opinion or binding decision from the \gls{edpb}, if it considers that final measures must be urgently adopted.\autocite[66(2)]{GDPR}
Alternatively, the \gls{edpb} can also issue urgent opinions or binding decisions on the request of an \gls{sa} even when it does not adopt measures, when the \gls{sa} believes there is an urgent need to act to protect the fundamental rights and freedoms of data subjects.\autocite[66(3)]{GDPR}

The urgency procedure can also be triggered following the failure of another \gls{sa} to comply with the cooperation procedure.
When an \gls{sa} fails to comply with a request for mutual assistance, the threshold for the urgency procedure is considered to be met, allowing an \gls{sa} to adopt provisional measures, as well as requesting an urgent binding decision from the \gls{edpb}.\autocite[61(8)]{GDPR}
Additionally, when an \gls{lsa} does not invite a \gls{sac} to participate in joint operations, or ignores a request of a \gls{sac} to participate within one month, the \gls{sac} may also adopt an urgent provisional measure, requiring an urgent binding decision of the \gls{edpb}.\autocite[62(7)]{GDPR}

\subsection{The Post-Reform Cooperation and Consistency Mechanism}

\todo[inline]{Add coverage of the information-exchange obligation (Art 9) and the cooperation file (Art 26) together --- these are the evidentiary-record provisions (LSA $\to$ SACs/EDPB). NB: ``cooperation file'' is already used below but never defined.}

As mentioned in Section 1.1.2, the Procedural Regulation only concerns cross-border processing.\autocite[1]{ProceduralRegulation}
This section highlights the amended flow of cooperation between the \gls{lsa} and the \glspl{sac} and other changes in cross-border enforcement. 
For the purposes of this section, a `receiving \gls{sa}' refers to the \gls{sa} receiving the initial complaint, before the allocation of competence is determined.
Additionally, the moment of confirmation of the competence of the \gls{lsa} is referred to as $T_0$.

\subsubsection{The Harmonization of `Parties Under Investigation' and complaint forms}

The Procedural Regulation provides a new definition for `parties under investigation', which refers to the controller or processor under investigation for an infringement of the \gls{gdpr} concerning cross border processing.\autocite[2]{ProceduralRegulation}

The Procedural Regulation also harmonizes the information to be contained in a complaint form, prohibiting an \gls{sa} from imposing additional requirements on the complainant.\autocite[4(1)]{ProceduralRegulation}
The harmonization only concerns complaints relating to cross-border processing.\autocite[4(1)]{ProceduralRegulation}
When a complainant does not include the required information, the receiving \gls{sa} must declare the complaint inadmissible and inform the complainant within two weeks of receipt.\autocite[4(2)]{ProceduralRegulation}

\subsubsection{The Proceduralization of Competence Attribution}

When an \gls{sa} receives a formally valid complaint, as outlined in Section 3.3.1, it must formulate a preliminary conclusion on: (i) whether the complaint concerns cross-border processing; (ii) the \gls{sa} it considers competent as \gls{lsa}; and (iii) whether the alleged infringement exclusively affects the \gls{ms} of its own competence, as explored in section 3.2.\autocite[4(4)]{ProceduralRegulation}

The receiving \gls{sa}  must, within six weeks from receipt, transmit the complaint to the \gls{lsa}.\autocite[4(5)]{ProceduralRegulation}
The \gls{lsa} will have six weeks to confirm its competence or contest it by referring the matter to the \gls{edpb} for dispute resolution.\autocite[4(6)]{ProceduralRegulation}
If the \gls{lsa} is inactive, at the expiration of the six weeks, the receiving \gls{sa} refers the matter to the \gls{edpb} for dispute resolution.\autocite[4(6)]{ProceduralRegulation}
The moment the competence is confirmed either by the LSA itself, or through a binding decision of the \gls{edpb} will be referred to as $T_0$.

\subsubsection{Three New Tracks for Cross-Border Enforcement Cases}

The Procedural Regulation provides for three tracks for the resolution of complaints relating to cross-border processing: early resolution, simple cooperation, and complex cooperation.

\paragraph{Early Resolution} 
Through early resolution, a complaint is dismissed when the alleged infringement has been brought to an end, making the complaint `devoid of purpose'.\autocite[5(2)]{ProceduralRegulation}
The procedure can be applied to cases that concern cross-border infringements relating to the rights of data subjects laid out in ch III of the \gls{gdpr}.\autocites[5(1)]{ProceduralRegulation}[ch III]{GDPR}
Early resolution can be triggered by the receiving \gls{sa} before it even transmits it to the \gls{lsa}, or by the \gls{lsa} after receiving the complaint.\autocite[5(1)]{ProceduralRegulation}
\todo[inline]{Reword this paragraph to include at "any time before the submission of the preliminary findings under Art. 19, or, where simple cooperation applies, before the submission of the draft decision."}

The \gls{sa} dismissing the complaint must establish, based on evidence, that the infringement has been brought to an end.\autocite[5(2)]{ProceduralRegulation}
When a complaint is dismissed through early resolution, the dismissing \gls{sa} must inform the complainant of the end of the infringement, that it considers the complaint `devoid of purpose', of the consequences of early resolution, and the complainant's right to object to early resolution within four weeks.\autocite[5(2)]{ProceduralRegulation}
When a complaint is dismissed by the receiving \gls{sa}, the receiving \gls{sa} establishes the resolution of the complaint within two weeks from the end of the objection period, and informs all parties.\autocite[5(3)]{ProceduralRegulation}
When a complaint is dismissed by the \gls{lsa}, the \gls{lsa} submits a draft decision to the \glspl{sac} within four weeks from the expiration of the objection period.\autocite[5(4)]{ProceduralRegulation}

It is important to note that when early resolution is employed, the procedure does not follow the new requirements provided for by the Procedural Regulation, including new procedural rights for complainants and parties under investigation.\autocite[5(6)]{ProceduralRegulation}

\paragraph{Simple Cooperation} 
A complaint can be handled through a simple cooperation procedure by the \gls{lsa} when, after having formed a preliminary view on the case, the \gls{lsa} believes that: 
\begin{enumerate*}[label=(\roman*)]
    \item there is no doubt regarding the scope of the investigation;
    \item and the case does not present legal or factual complexity that would require additional cooperation with the \glspl{sac}.
\end{enumerate*}\autocite[6(1)]{ProceduralRegulation}
Simple cooperation should not be employed in cases where there is a systemic risk or recurring problems in multiple \glspl{ms}, when the case intersects with other legal fields other than data protection, or when the case affects a large number of data subjects or is based on multiple complaints across the Union.\autocite[recital 26]{ProceduralRegulation}

To trigger the simple cooperation procedure, the \gls{lsa} must inform the \glspl{sac} of its intention to handle the case through simple cooperation within six weeks from $T_0$.\autocite[6(2)]{ProceduralRegulation}
Any \gls{sac} may object to the application of the simple cooperation procedure within two weeks of the notification, and the procedure will continue according to the new complex procedure.\autocite[6(3)]{ProceduralRegulation}

Under the simple procedure, an \gls{lsa} must deliver a draft decision to the \glspl{sac} within twelve months of $T_0$.\autocite[12(6)]{ProceduralRegulation}
An LSA may extend the deadline by no more than two months when national law requires steps to be taken by the LSA that cannot be completed within the twelve months.\autocite[12(6)]{ProceduralRegulation}
Aside from that, most the new procedural requirements and time-limits established by the Procedural Regulation do not apply to the simple procedure.\autocite[6(1)]{ProceduralRegulation}
Importantly, this excludes new rights of the parties under investigation and complainants to be heard, with the simple cooperation procedure stating that the parties should have a right to be heard, but leaving it to the \glspl{ms} to determine the procedural rights of the parties.\autocite[6(4)]{ProceduralRegulation}

\paragraph{Complex Cooperation}
The biggest changes introduced by the Procedural Regulation concern a new cooperation flow, referred to as `complex procedure'. 
The complex cooperation procedure presents many more procedural steps to the cooperation and consistency mechanism, which will be explored in detail next.

It must be noted at the outset, that the Procedural Regulation specifically states that the new procedural steps and time limits do not influence the rights of the parties under investigation and complainants,\autocite[7]{ProceduralRegulation} nor that missing a time limit will influence the validity or legality of a decision.\autocite[14]{ProceduralRegulation}

At start of an investigation, after forming a preliminary view on a case, an \gls{lsa} must now draft a summary of key issues and present it to the \glspl{sac} within three months from $T_0$.\autocite[10(1), (3)]{ProceduralRegulation}
The summary of key issues must include:
\begin{enumerate*}[label=(\roman*)]
    \item the main relevant facts;
    \item the scope of the investigation, including the legal provisions allegedly infringed;
    \item where available, an analysis of the views of the parties to the complaint
    \item where applicable, a preliminary identification of potential corrective measures.
\end{enumerate*}\autocite[10(2)]{ProceduralRegulation}
The receiving \gls{sa} may collaborate with the \gls{lsa} by providing relevant information for the summary to the \gls{lsa} when transferring the complaint.\autocite[10(6)]{ProceduralRegulation}
\glspl{sac} will have four weeks -- with a possibility for a two-week extension based on the complexity of the case or on their request -- to review the summary and provide comments.\autocite[10(4)]{ProceduralRegulation}
\glspl{sa} must share all comments with each other.\autocite[10(5)]{ProceduralRegulation}
The \gls{lsa}, within four weeks -- or six, depending on the complexity of the case -- from the deadline for the submission of comments by \glspl{sac}, must respond to the comments and outline whether and how it plans to take them into account.\autocite[10(5)]{ProceduralRegulation}

The Procedural Regulation reiterates that all \glspl{sa} must cooperate in an endeavor to reach consensus, and that they may use all the tools provided by the cooperation procedure laid out in the \gls{gdpr} to do so.\autocite[11(1)]{ProceduralRegulation}
The Procedural Regulation however introduces a new procedure to find consensus at the earlier stages of a cross-border investigation.
When a \gls{sac} disagrees with the \gls{lsa} and consensus cannot be found, the \gls{sac} can, within a month from the deadline for the \gls{lsa} to respond to the comments on the summary of key issues,\autocite[11(3)]{ProceduralRegulation} request mutual assistance or to carry out joint operations in order to reach consensus on: 
\begin{enumerate*}[label=(\roman*)]
    \item the scope of the investigation and the points of law;
    \item the legal or factual issues;
    \item and the corrective measures to be adopted in case an infringement is found.
\end{enumerate*}\autocite[11(2)]{ProceduralRegulation}
The \gls{lsa} must now respond to either request within one month of its receipt.\autocite[11(4)]{ProceduralRegulation}
If consensus cannot be found on the scope of the investigation and the points of law to be investigated, the \gls{lsa} must refer the matter to the \gls{edpb} requesting an urgent binding decision.\autocite[11(6)]{ProceduralRegulation}

\todo[inline]{connect to the binding decisions and the opinion}
In the decisions presented, \gls{edpb} called on the \glspl{sa} to reach consensus on the legal and factual scope of the investigations at the earlier stages of the procedure

\subsubsection{The Draft Decision}

The Procedural Regulation spells out more detailed procedures for the adoption of a final decision on a case, especially regarding the new complex procedure.
The \gls{lsa} must submit a draft decision pursuant to Article 60(3) \gls{gdpr} to the \glspl{sac} within fifteen months from $T_0$.\autocite[12(1)]{ProceduralRegulation}
That time limit is suspended in case the \gls{lsa} requests an urgent binding opinion from the \gls{edpb} on the lack of consensus between \glspl{sa}, as explored earlier.\autocite[12(2)]{ProceduralRegulation}
The \gls{lsa} may also extend the fifteen-month deadline by a maximum of twelve months on an exceptional basis.\autocite[12(3)]{ProceduralRegulation}
It must do so no later than four weeks from the expiration of the fifteen-month deadline, informing the other \glspl{sa} of the reasons and of the length of the extension.\autocite[12(3)]{ProceduralRegulation}

The \glspl{sac} may object to the extension of the deadline with a reasoned request submitted within two weeks of the receipt of the notification of the extension by the \gls{lsa}.\autocite[12(4)]{ProceduralRegulation}
The \gls{lsa} must take into account the objection.\autocite[12(3)]{ProceduralRegulation} 
However, it may still extend the deadline.\autocite[12(5)]{ProceduralRegulation}
In that case, the \glspl{sac} may inform the \gls{lsa} that there is a need to act in order to protect the right and freedoms of data subjects.\autocite[12(5)]{ProceduralRegulation}
If the \gls{lsa} does not submit a draft decision within the extended time period, the \glspl{sac} that informed the \gls{lsa} of the need to act may adopt provisional measures on their own territory, activating the urgency procedure in Article 66(1) \gls{gdpr}.\autocite[12(5)]{ProceduralRegulation}

\paragraph{(Partial) Dismissal of Complaints} 
When an \gls{lsa} plans to submit draft decision dismissing a complaint in whole or in part, it must first inform the receiving \gls{sa} of its preliminary view on the dismissal, before the draft decision is presented.\autocite[16(1)]{ProceduralRegulation}
The receiving \gls{sa} must then inform the complainant and provide them with an opportunity to make their views known in writing, with a time limit between three and six weeks.\autocite[16(1)]{ProceduralRegulation}
The receiving \gls{sa} must then share the complainant's views with the \gls{lsa} no later than one week from receipt.\autocite[16(2)]{ProceduralRegulation}
If the complainant's views do not change the preliminary assessment of the \gls{lsa}, the \gls{lsa} can proceed with the submission of the draft decision.\autocite[16(3)]{ProceduralRegulation}
When a complaint is (partly) rejected, the receiving \gls{sa} must inform the complainant of the judicial remedy available to them.\autocite[18]{ProceduralRegulation}

\subsubsection{The Rights of Parties Under Investigation and of Complainants}

The Procedural Regulation introduces some procedural rights for parties under investigation and complainants in cross-border enforcement cases. 
These are especially relevant in the context of the new complex procedure.

Before drafting a draft decision to be sent to the \glspl{sac}, the \gls{lsa} now needs to draft preliminary findings on the case.\autocite[19(1)]{ProceduralRegulation}
Preliminary findings must include the findings of the investigation and the allegations raised to enable the parties under investigation to understand the case that is being raised against them.\autocite[19(2)]{ProceduralRegulation}
The findings must include the corrective measures the \gls{lsa} is considering to adopt, based on the information available at the time of drafting.\autocite[19(2)]{ProceduralRegulation}
The preliminary findings must be shared with the \glspl{sac}, who will have four weeks -- or six -- to provide comments.\autocite[19(3)]{ProceduralRegulation}

The preliminary findings must be shared then, by the \gls{lsa}, with the parties under investigation, and by the receiving \gls{sa} with the complainant.\autocite[19(4) and 20(1)]{ProceduralRegulation}

\paragraph{Parties Under Investigation} The Procedural Regulation highlights that parties under investigation and complainants are not in the same procedural position, and should not be treated as such.\autocite[recital 50]{ProceduralRegulation}

The parties under investigation must receive the preliminary observations by the \gls{lsa}, and be given the opportunity to make their views known in writing or orally at a hearing between 3 and 6 weeks from their receipt.\autocite[19(5)]{ProceduralRegulation}

Importantly, the facts, allegations, and legal assessments included in the preliminary findings become binding for the \gls{lsa}.
The \gls{lsa} is only allowed to base its draft decision on the findings for which the parties under investigation have had the opportunity to present their views on.\autocite[19(7)]{ProceduralRegulation}

Additionally, the parties under investigation have the right to make their views heard when a revised draft decision raises new allegations or legal assessments that were not included in the preliminary findings.\autocite[22(1)]{ProceduralRegulation}
These must be requested by the \gls{lsa} prior to submitting a revised draft decision to the \glspl{sac}.\autocite[22(1)]{ProceduralRegulation}

\paragraph{Complainants} Complainants too must receive the preliminary findings by the receiving \gls{sa}, and must be given the opportunity to make their views known to the \gls{lsa} only in writing within three to six weeks from their receipt.\autocite[20(1)]{ProceduralRegulation}
However, the Procedural Regulation states that national law still governs the procedural rights of complainants at this stage.\autocite[20(2)]{ProceduralRegulation}

\subsubsection{Relevant and Reasoned Objections}

Once the \gls{lsa} submits a draft decision, the \glspl{sac} retain the right to submit relevant and reasoned objections according to Article 60(4) and 60(5) \gls{gdpr}.
The Procedural Regulation provides new rules on the content of relevant and reasoned objections, especially in the context of the new complex procedure.
A relevant and reasoned objection must be based on factual or legal elements found in the draft decision or the cooperation file, they cannot concern the scope of an investigation where consensus was reached beforehand, and they cannot concern a decision made with an early resolution procedure.\autocite[23(1)]{ProceduralRegulation}

\glspl{sac} may however object to the scope of the investigation at the draft decision stage if the \gls{lsa} failed to investigate all the elements on which consensus was reached, or if new elements that pose a risk for the fundamental rights and freedoms of data subjects have surfaced since the investigation.\autocite[23(2)]{ProceduralRegulation}

The relevant and reasoned objection must be sufficiently clear, coherent and precise to enable the \glspl{sa} to prepare their position, and eventually to enable the \gls{edpb} to efficiently resolve a potential dispute.\autocite[23(3)]{ProceduralRegulation}

\subsubsection{Adoption of the Final Decision}

Within a month from the end of the period for submission of comments on the draft decision, the \gls{lsa} must adopt the decision as established in Articles 60(7) or 60(9) \gls{gdpr}, and inform the main establishment of the controller or processor found in breach.\autocite[21(1)]{ProceduralRegulation}

The receiving \gls{sa} must then inform the complainant of the decision, and provide them with a summary of the decision, as well as the full decision, excluding elements that are confidential.\autocite[21(2)]{ProceduralRegulation}

\subsubsection{Changes to Dispute Resolution and Urgency Procedures}

\paragraph{Dispute Resolution}

When revising a draft decision, the \gls{lsa} must now present the revised draft to the \glspl{sac} within three months from the end of the period for submission of comments, or refer the matter to the \gls{edpb} for dispute resolution.
\autocite[27(1)--(2)]{ProceduralRegulation}
That referral must contain the draft decision, the relevant facts, the views of the parties under investigation and complainants, the objections made by the \glspl{sac}, and the reasons of the \gls{lsa} for not following the objections.\autocite[27(3)]{ProceduralRegulation}
The \gls{edpb} must also now provide a preliminary view on whether the objections are relevant and reasoned within four weeks of referral.\autocite[27(5)]{ProceduralRegulation}
In this case, the \gls{edpb} must also now invite the parties under investigation, or -- if the \gls{edpb} intends to discard (part of) a complaint -- the complainant, to provide their views in writing within two weeks when its decision will be based on new factual or legal elements.
\autocite[28(1)]{ProceduralRegulation}

The Procedural Regulation also lays down detailed requirements on dispute resolution for \gls{lsa} competence attribution, and for the resolution of disputes regarding the failure of an \gls{sa} to request or follow the opinion of the \gls{edpb}.\autocite[29--30]{ProceduralRegulation}
Regarding the latter, the \gls{sa} referring the matter must now substantiate where and how the failure to comply with the requirements in Article 64 \gls{gdpr} has happened. \autocite[30(1)]{ProceduralRegulation}
The \gls{edpb} must then ask the \gls{sa} alleged to be in breach to provide its views.\autocite[30(2)]{ProceduralRegulation}

\paragraph{Urgency Procedure}

The Procedural Regulation now contains more detailed procedures for requesting an urgent opinion or binding decision from the \gls{edpb} under Article 66 \gls{gdpr}.

A request for an urgent opinion under Article 66(2) \gls{gdpr} must be filed four weeks before the end of the provisional measures at the latest, and the \gls{sa} must now justify the urgent need to act, and explain the exceptional circumstances that require the request for an urgent opinion.\autocite[31(1)]{ProceduralRegulation}

When requesting an urgent binding decision under Article 66(2) \gls{gdpr}, the request must also contain, in addition to the requirements for an urgent opinion, the investigative measures taken by the \gls{sa}, the views of the \gls{lsa} and the views of the local establishment of the controller or processor.\autocite[32(1)]{ProceduralRegulation}

The Procedural Regulation also provides for new requirements for requesting an urgent opinion or binding decision under Article 66(3) \gls{gdpr}.\autocite[33(1)]{ProceduralRegulation}