\section{Safeguarding the Right to Data Protection as a Fundamental Right}

Good administration provides the basis for efficient, fair, and rational administrative decision making', as well as guaranteeing individual procedural rights.\autocite[213]{hofmannDemkovaPrinciplesProcJustice}
Compliance with the standards set by good administration upholds the integrity of the public administrative system, as well as an individual's understanding of the reasons and input of a decision, enabling them to understand whether their rights have been violated and seek redress if necessary.\autocite[213]{hofmannDemkovaPrinciplesProcJustice}

This study examined the reforms introduced by the Procedural Regulation insofar as they relate to the functioning of the \gls{gdpr}'s cooperation mechanism, a network of peer authorities that are tasked with a uniform application of the law protecting the fundamental right to data protection.
So far, this thesis focused on the role of the actors involved in cross-border enforcement of the \gls{gdpr}, and their ability to push each other to comply with the standards of good administration through the tools of administrative cooperation provided in data protection law.
This chapter will consider whether good administration, as a principle, is an effective tool for safeguarding the right to data protection as a fundamental right, and what consequences stem from a failure to comply with its standards in the context of that fundamental right.

\subsection{Process Rights and Effective Protection for Individuals}

The institution of procedural rights is deeply connected to the respect of human dignity.\autocite[75--76]{galliganProceduralFairness1997}
Fair procedures bring fair treatment, and fair treatment is a necessity if we asssume that individuals are entitled to dignity and respect, especially in a community that is based on the rule of law.\autocites[75--76]{galliganProceduralFairness1997}[338--339]{nehlGoodAdminProceduralRight2009}
Procedural guarantees may not always be enough to ensure effective protection for the individual and collective, but function as a framework to assess the powers of the public authority, and seek redress when those powers are abused.\autocite[76--77]{galliganProceduralFairness1997}

In this context, good administration provides the judiciary with a flexible framework to calibrate the intensity of judicial oversight.\autocite[102]{hofmannDutyCareEU2020}
Judicial oversight, however, is limited by the separation of powers.\autocite[4]{mendesDiscretionCarePublic2016}
Review is not a substitute for administrative discretion, and the judiciary cannot command the choices made within the adminsitration's discretionary power.\autocite[4]{mendesDiscretionCarePublic2016}

\subsection{Why is Data Protection a Fundamental Right Anyway?}

Data protection law was developed in response to the advancements in computational technology of the 1970s and the fear that databases could enable surveillance at a large scale.\autocite[905]{streinzEvolutionEuropeanData2021}
As Daniel Solove argues -- using Kafka's \emph{The Trial} as a metaphor -- data processing at large scale strips individuals of their autonomy.\autocite[1420]{solovePrivacyPowerComputer2000}
In the novel, the protagonist, is arrested and put on trial by a mysterious, large authority who had ammassed a vast dossier on him, but never informed him of the reasons for his arrest.\autocite[1420]{solovePrivacyPowerComputer2000}
The novel illustrates the helplessness, frustration, and vulnerability of an individual who is dehumanized, reduced to a vast dossier of his data, that he is not aware of, has no control upon, and is unable to challenge.\autocite[1420]{solovePrivacyPowerComputer2000}

The right to data protection, as mentioned throughout this study, is a fundamental right enshrined in \gls{eu} primary law.\autocites[8]{CFR}[16]{TFEU}
Its implementation in secondary law, the \gls{gdpr}, provides an intricate framework for the protection of personal data, which ensures that data processing is lawful, limited to what is necessary for the purposes for which it is processed, enabling individuals to understand how, why and by who the processing of their personal data is carried out.

The right to data protection also requires the enforcement of said right to be carried out by an independent authority.
Enforcement, is not just an obligation under secondary law, but a constitutional duty. 

\subsection{Is Good Administration an Effective Tool for Safeguarding the Right to Data Protection?}

The principles examined in this study -- namely the duty of the administration to deliver careful and reasoned decisions -- if violated, may form grounds for the annullment of the administrative decision.\autocites[29]{TUM}[78]{Sytraval}
Annulment, however, is not a substitute for the procedures that are in place to enforce the law, which in the case of the \gls{gdpr} are the cooperation and consistency mechanism.
Futhermore, annulment forms an individual remedy, that needs to be sought out by the individual, and will set aside only the contested decision.

While good administration has served as a benchmark for the assessment of the amended cross-border enforcement system provided by the Procedural Regulation, it is not a self-executing tool for ensuring the protection of the fundamental right to data protection.

\subsection{The Consequences of a Failure to Enforce the Right to Data Protection}

As seen above, the right to data protection provides us with a right to informational self-determination, especially in light of the consequences the processing of personal data may bring in relation fundamental freedoms, such as the right to private life.\autocites[68]{ÖsterreichischerRundfunk}[7]{CFR}[8]{echr}
Data protection is not an absolute right,\autocite[recital 4]{GDPR} but processing must always be carried out in accordance with the law,\autocite[8(2)]{CFR} to ensure that the balance between the protection of fundamental rights and the interests requiring the free movement of data is fair.\autocite[42]{SchremsI}
The guarantee of independence of the supervisory authority is intended to ensure the effectiveness and reliability of the system of data protection, and supervision is therefore an essential component of the protection of individuals against infringements of their fundamental right to data protection.\autocite[41]{SchremsI}

Striking a balance between the protection of fundamental rights and the free movement of data is no easy feat.
Advancements in technology -- and in particular more recently, the use of neural networks -- have enabled the processing of personal data at a scale unimaginable even a decade ago.\autocite[recital 6]{GDPR}
For this reason, \gls{gdpr} created a system of principles and rules that are meant to be technologically neutral, to ensure the law cannot be circumvented.\autocite[recital 15]{GDPR}

It is outside the scope of this study to assess whether the secondary-law interpretation of the fundamental right to data protection is adequeate to protect that right.
It is crucial, however, that the enforcement of that law is carried out consistently, predictably, carefully, to ensure that the fundamental right to data protection is safeguarded with respect to the processing that is happening today, and in the years that lie ahead.
Many new pieces of legislation refer to the \gls{gdpr}, or are without prejudice to it, requiring a careful examination of the interactions with more specific legislation, and how the obligations under the \gls{gdpr} will shape the emerging digital acquis.
 
This study concluded that \glspl{dpa} must be careful and base their decisions on reasoned arguments, as a prerequisite under their consitutional duty to respect the principle of good administration.
In the context of cross-border enforcement, as recognized by AG Bobek, a disjointed system of oversight creates `hawks and doves', leading to forum shopping and an evasion of the law.
The dominance of the \gls{lsa} in setting the informational and legal scope of cross-border investigations has generated manifest failures to investigate matters to their fullest, thus undermining the respect of the principle of good administration, and ultimately the protection of the fundamental right to data protection.

The Procedural Regulation provides an intricate flowchart of procedural steps meant to provide more clarity on the procedures to be followed in cross-border cases.
It does not, however, provide penalties for failing to comply with the new procedural rules, and in some cases, it even gives \glspl{dpa} the discretion to dismiss complaints.
The Procedural Regulation does not address the deeper problems of the integrated system of enforcement of the \gls{gdpr}, such as chronic underfunding of \glspl{dpa}, or the lack of harmonization of administrative procedure in the \gls{eu}, which impacts the uniformity of enforcement procedures across the \gls{eu}.