\section{Introduction}

\subsection{Background, Problem, and Analytical Framework}

On 26 November 2025, Regulation 2025/2518 laying down additional procedural rules on the enforcement of Regulation (EU) 2016/679\autocite{ProceduralRegulation} (hereinafter `Procedural Regulation') was adopted. 
The Procedural Regulation seeks to further proceduralize the cooperation and consistency mechanism of \gls{gdpr}\autocite{GDPR} in cases involving cross-border enforcement.

The \gls{gdpr} is the cornerstone of the \glsdisp{eu}{European Union's (EU)} so-called `digital acquis'.\autocite[223]{zanfirGDPREDPS} 
Entered into force in 2018, and replacing the Data Protection Directive of 1995, it has often been heralded as the most comprehensive package of data protection and privacy regulations.\autocite[498]{mustertProceduralisingEnforcement2026} 
The \gls{gdpr} operationalizes the fundamental right to data protection, as enshrined in primary \gls{eu} Law.\autocites[8]{CFR}[16]{TFEU}

The \gls{gdpr} seeks to provide a harmonized and directly applicable legal framework for the processing of personal data across the \gls{eu}, increasing legal certainty and strengthening enforcement.\autocite[recital 7]{GDPR}
However, the \gls{gdpr}'s enforcement mechanism has encountered practical challenges, particularly in the context of cross-border enforcement.

\subsubsection{The Enforcement Challenges of the GDPR}

The \gls{gdpr} relies on a system of decentralized enforcement.\autocite[recital 1]{ProceduralRegulation} 
\glspl{ms} are to establish independent \glspl{dpa}, who are competent to carry out enforcement in their national territory.\autocite[51(1) and (2)]{GDPR}

Cross-border enforcement is handled through a \gls{oss} system,\autocite[recital 127]{GDPR} through the cooperation and consistency mechanism.\autocite[chapter VII]{GDPR} 
When an alleged infringement concerns processing taking place in more than one \gls{ms}, the \gls{dpa} of the \gls{ms} where the main establishment of the controller or processor is located assumes the role of \gls{lsa}.\autocite[56(1)]{GDPR}
The \gls{lsa} acts as the sole interlocutor of the accused.\autocite[63]{facebookBelgium}
Other \glspl{dpa} with a vested interest in the case become \glspl{sac}, and are to cooperate with the \gls{lsa} in investigating the infringement.\autocite[60]{GDPR}
In case cooperation fails, the consistency mechanism is triggered,  involving the \gls{edpb} as a settler of disputes between the \gls{lsa} and the \glspl{sac}.\autocite[arts 63--66]{GDPR}

Multiple scholars and stakeholders of data protection law have noted that the \gls{gdpr}'s enforcement design has suffered from numerous flaws.
Institutional actors and scholars have signaled a lack of true cooperation between \glspl{dpa}, due to differences in administrative procedure,\autocites[324--325]{mustertCrossborderEnforcementGDPR}[1, 3, 4]{EDPBLetterEUComm} diverging national priorities,\autocite[806]{gentileDeficientDesign2022} and weak procedural rights for complainants in cross-border cases.\autocite[800]{gentileDeficientDesign2022} 

Another concern rests on the role the \gls{lsa} plays in the cooperation procedure. 
The cooperation procedure requires the \glspl{sa} to cooperate through information exchange and mutual assistance in an endeavor to reach consensus.\autocites[60(1) and (2)]{GDPR}[point 111]{fbBelgiumAgBobek}
In practice, the \gls{lsa} has been enabled to exert significant influence over the scope of an investigation, the determination of corrective measures, and the enforcement of those measures.\autocite[809]{gentileDeficientDesign2022}

\subsubsection{The Procedural Regulation}

Against this backdrop, in July 2023 the Commission presented a proposal for a Procedural Regulation.\autocite{ProposalProcReg} 
The Procedural Regulation was adopted in November 2025.
In its final version, the Procedural Regulation harmonizes aspects relating to the admissibility\autocite[4]{ProceduralRegulation} and early resolution of complaints,\autocite[5]{ProceduralRegulation} the further proceduralization of administrative cooperation under Article 60 \gls{gdpr} (including the institution of time limits),\autocite[ch III]{ProceduralRegulation} and the procedural rights of `parties under investigation' and of complainants.\autocite[art 2 and s 5]{ProceduralRegulation}

The Procedural Regulation only applies to infringements with a cross-border element,\autocite[499]{mustertProceduralisingEnforcement2026} leaving unaddressed many of the concerns on the lack of harmonization of national administrative procedure.\autocite[499 and 506]{mustertProceduralisingEnforcement2026}
The Procedural Regulation however seems to act in particular on one aspect of the enforcement design of the \gls{gdpr}: the discretion afforded to the \gls{lsa} in the cooperation procedure. 

\subsubsection{Peer Pressuring the LSA into Action?}

While the \gls{lsa} is the chief actor in a cross-border case, all \glspl{sa} are collectively responsible for the protection of the rights afforded by the \gls{gdpr}.\autocite[paras 63--68]{facebookBelgium}

This framing may be viewed as a form of `peer pressure', where enforcement is pushed forward by horizontal cooperation.\autocite[466]{hofmannmustertScholtenHandbook2023} 
The architecture of the \gls{gdpr} can be described as a network of horizontal collaboration (between the \glspl{dpa}), with a vertical element of norm-setting and decision-making in the \gls{edpb}.\autocite[428]{mastenbroekFillingGapEuropean2018}

While the Procedural Regulation does not expand individual rights significantly, the efficiency of the enforcement procedure is fundamentally connected to the protection of Europeans' fundamental rights, the \gls{gdpr} being the secondary law implementation of the fundamental right to data protection.\autocite[16]{TFEU}
To evaluate the potential for effectiveness of the reform, this thesis will assess the Procedural Regulation against the general principle of good administration.

\subsubsection{The Principle of Good Administration as a Benchmark}

The principle of good administration is one of the general principles of \gls{eu} law,\autocite[49]{HNvMinisterJustice} as well as being a \gls{cfr} right for individuals vis-à-vis EU institutions.\autocite[41]{CFR}

The principle of good administration encompasses a number of procedural rights and guarantees in the realm of administrative proceedings.
It presents two facets: an objective and a subjective rationale.\autocite[111--112]{mihaescu-evansRightGoodAdministration2015}
The objective rationale ensures the efficiency of the administration, while the subjective rationale aims to provide individuals with procedural guarantees when facing an administrative authority.\autocite[111--112]{mihaescu-evansRightGoodAdministration2015}

With the explicit aim of bolstering efficiency in handling cross-border enforcement,\autocite[recital 17]{ProceduralRegulation} the Procedural Regulation's amended cooperation procedure may bring \glspl{dpa} closer to the requirements of the principle of good administration. 

\subsubsection{The Duty of Care}
The duty of care, as a sub-principle of good administration, requires administrations to act diligently and carefully when making decisions\autocite[92--94]{hofmannDutyCareEU2020}, and to conduct a `full and impartial investigation of a fact set prior to decision-making [...]'.\autocite[84]{hofmannEvansRelationCharterFundamental2013}
In particular, a requirement for early consensus building among \glspl{dpa}\autocite[11]{ProceduralRegulation} may contribute to bolstering the respect of the duty of care by \glspl{sa}. 

By requiring \glspl{dpa} to reach consensus on the scope and the points of law the investigation will cover, the reform may bring a more diligent and thorough investigation, as well as more consistent enforcement.

\subsubsection{The Duty to Give Reasons}

The renewed multi-step cooperation procedure highlights a vital need to substantiate the logic behind any decision or assessment while collaborating on a case. 
Often, European enforcement networks are built around information sharing.\autocite[140]{hofmannCompositeDecisionMaking2009} 
Enforcement under the \gls{gdpr} is no exception, with mutual assistance forming an integral part of the enforcement design of the \gls{gdpr}.\autocite[61]{GDPR} 

The duty to give reasons functions as a check on the administration's conduct before it reaches a decision.\autocite[369--370]{craigProcess2018}
The exercise of this duty by the administration may help it to identify possible flaws in its reasoning, by having to explain them.\autocite[369--370]{craigProcess2018}

Mandatory structured cooperation from an early stage may push \glspl{dpa} to share substantiated assessments of the legal and factual aspects of a case. 
This becomes all the more important in the context of dispute settlement by the \gls{edpb}, who relies on the information provided by the \glspl{dpa} to reach a binding decision.

\subsection{Research Gap}
The Procedural Regulation will enter into force on 2 April 2027.\autocite[37]{ProceduralRegulation}
Mustert and Hofmann provided early commentary on the Commission Proposal in 2024 while it was in the trilogue stage,\autocite{hofmannFutureGDPREnforcement2024} and Mustert provided an early comment on the Regulation after its passing.\autocite{mustertProceduralisingEnforcement2026} 
However, no structured assessments of the reformed cooperation procedure against \gls{eu} primary law have been conducted. 
This thesis will attempt to conduct such an assessment, following Mustert's example,\autocite{mustertCrossborderEnforcementGDPR} albeit at smaller scale.

\subsection{Research Questions}
\begin{itemize}
    \item To what extent does Regulation 2025/2518 curb \glspl{lsa}' discretionary powers in light of the duties of care and to give reasons under the general principle of good administration in the context of composite cross-border enforcement of the \gls{gdpr}?
    \begin{itemize}
        \item What do the duty of care and duty to give reasons require of \glspl{dpa} in the context of cross-border \gls{gdpr} enforcement?
        \item What changes does the Procedural Regulation bring to the \gls{gdpr}'s cooperation and consistency mechanism?
        \item Are earlier consensus finding and new procedural deadlines under Reg 2025/2518 adequate to fulfill the duty of care of \glspl{dpa}?
        \item Will earlier consensus finding and new procedural deadlines contribute to respecting the duty to give reasons in composite administrative proceedings involving the \gls{edpb}?
        \item What consequences stem from a failure to observe the principle of good administration for the protection of fundamental rights in the context of cross-border enforcement of the \gls{gdpr}?
    \end{itemize}
\end{itemize}

\subsection{Methodology}

\subsubsection{Methods}

To answer the research questions, the thesis will employ a primarily doctrinal research, accompanied by desk research. 
Doctrinal research aims to give a systematic account of the principles, rules and concepts that govern a legal field.\autocite[210]{smitsWhatLegalDoctrine2017}
It analyzes the relationship between those principles to answer gaps in the existing law.\autocite[210]{smitsWhatLegalDoctrine2017}

In the context of this thesis, doctrinal research will be employed to examine the developments to the cooperation and consistency mechanism of the \gls{gdpr} provided by the Procedural Regulation. 
This will be complemented with the jurisprudence of the \gls{cjeu} in the field of \gls{gdpr} enforcement by \glspl{dpa}, \gls{edpb} guidelines, \gls{edps} documents, and other official documents relating to \gls{gdpr} enforcement and the Procedural Regulation. 
This method will be necessary to map the changes to the cross-border enforcement of the \gls{gdpr} as amended by the Procedural Regulation, and to ascertain the duties of \glspl{dpa} in enforcing data protection law. 

To ascertain the duties of administrations under the principle of good administration, a doctrinal analysis of the case law of the \gls{cjeu} on a reasoned and careful decision will be conducted.

The doctrinal research will also be complemented by relevant desk research looking into the literature on the enforcement of the \gls{gdpr} and the principle of good administration.\autocite[223]{smitsWhatLegalDoctrine2017} 
Secondary sources such as journal articles, books, PhD dissertations and other academic sources will be employed. 

To search for literature, online search engines such as Google Scholar, LexisNexis, and WorldCat will be used. 
Boolean search operators will be employed, for example:
\begin{itemize}
    \item (``GDPR'' OR ``General Data Protection Regulation'')
    \item (``cooperation'' OR ``one-stop-shop'' OR ``consistency mechanism'')
    \item (``enforcement'' OR ``supervisory authority'' OR ``lead authority'')
    \item (``procedural regulation'' OR ``Regulation 2025/2518'' OR ``COM(2023) 348'')
\end{itemize}

\subsubsection{Limitations}

This research is inspired by Mustert's comprehensive analysis of the \gls{gdpr} enforcement system against the constitutional foundations of the \gls{eu}.\autocite{mustertCrossborderEnforcementGDPR}
However, this study will look at the impact of the Procedural Regulation on enforcement only through the lens of two of the sub-principles of good administration and cannot replicate the breadth of the analysis Mustert conducted.
These principles have been selected due to a preliminary interpretation of the most impactful reforms to the \gls{gdpr} enforcement procedure. 
Many of the concerns Mustert has raised do not seem to find an answer in this Procedural Regulation, while others are only partially addressed.\autocite[]{mustertProceduralisingEnforcement2026}
This thesis however will only explore the new cross-border enforcement procedure. 

\subsubsection{Use of Generative AI}

Generative AI will be used for minor tasks such as brainstorming, proofreading, and helping with surfacing relevant points in the literature to explore. 
Anthropic's Claude Sonnet and Opus models will be employed, with their thinking mode enabled. 
Claude will be used with a Pro subscription inside of Claude Code\autocite{HowClaudeCode} and Cowork.\autocite{GetStartedCowork}

Claude Code and Cowork allow the user to position the program, running locally, within a defined folder.\autocite{HowClaudeCode}
These tools allow for the model to have specific instructions on how to operate, giving it context into the precise work of the user that's being conducted within that folder.\autocite{HowClaudeRemembers}
This is done through a CLAUDE.md file.\autocite{HowClaudeRemembers}
These instructions constrain the model into acting in specific ways, in light of the guidelines for Responsible AI use, making Claude act as a brainstorm companion.
The CLAUDE.md file used during the writing process will be attached in the AI logbook.

Claude will also be able to read sources directly from Zotero through a \gls{mcp} server.\autocite{yu54yyyuZoteromcp2026}
An \gls{mcp} server is an open-source standard that allows AI tools to integrate databases and other tools in a structured way.\autocite{WhatModelContext}
Claude will be able to access notes and annotations done within Zotero, as well as the full text of the sources saved in the library.
This will help ground the model into the correct sources and limit hallucinations.

\subsection{Outline}
\todo[inline]{Supervisor: To be added.}

The thesis will be structured as follows:
\begin{itemize}
    \item Chapter 2 will provide an overview of the duty to give reasons and the duty of care under the principle of good administration.
    \item Chapter 3 will provide an overview of the Procedural Regulation and its changes to the \gls{gdpr}'s cooperation and consistency mechanism.
    \item Chapter 4 will analyze the impact of the Procedural Regulation on DPAs ability to comply with the duty of care.
    \item Chapter 5 will analyze the impact of the Procedural Regulation on DPAs ability to comply with the duty to give reasons.
    \item Chapter 6 will analyze the consequences of a failure to observe the principle of good administration for the protection of fundamental rights
\end{itemize}